The impact of gas bulk rotation on the Lyman-$\alpha$ line

### 2014-02-13 paper/ms.tex
figures
spelling

The codes follow the individual scatterings of \ly photons as it
travels through a 3D distribution of neutral Hydrogen. At each
scattering the frequency of the photon (in the laboratory frame) and
its direction of propagation change. This change in frequency is due
to the peculiar velocities of the Hydrogen atom that absorbed and
reemitted the photon. If dust is present, the photon can interact
either with a Hydrogen atom or dust grain. In the case of a dust
interaction the photon can be either absorbed or scattered, this
probability is encoded in the dust albedo, $A$, which we chose to be
$1/2$. In order to obtain accurate values for the escape fraction of
photons in the presence of dust, we do not use any accelerating
mechanism in the radiative transfer. Once the photons escape the gas
distribution we store their direction at their direction of
propagation and frequency at their last scattering.

The photons are thus emitted in some region of the gas distribution
and follow a random walk in space and frequency until they escape the
gas distribution or are absorbed by a dust grain. The initialization
process for the \ly photons has to specify its position, frequency and
direction of propagation. In our case we select the initial frequency
to be exactly the \ly restframe frequency ($x=0$) and the direction of
propagation to be random following an flat probability distribution
over the sphere.

@@

\begin{figure*}
\begin{center}
  \includegraphics[width=1.0\textwidth]{f1.eps}
\end{center}
\caption{Shape of the \ly line for
    different velocities rotational velocities for spherical
    distributions in all the models. The continous (dashed) line represents the 
    central (homogeneous) source distributions.
    \label{fig:differentvelocities}}  
\end{figure*}

@@

In the next subsections we characterize the line morphology by the
full width at half maximum (FWHM) and the location of the peak
maxima. In order to interpret the morphological changes in the line we
also report the median number of scatter for each \ly photon in the
simulation. For the models where dust is included we measure the 
escape fraction as a function of rotational velocity. Finally, we make
an estimate of the anisotriopic emission of the models in comparison
with static spheres.

@@

\begin{figure*}
\begin{center}
  \includegraphics[width=0.9\textwidth]{f2.eps}
\end{center}
  \caption{FWHM for the non-dusty models as a function of
    rotational velocity $V_{\rm max}$. The left panel shows the
    results in velocity units while the right panel normalizes
    the data by the FWHM in the static case. 
    Continuous (dashed) lines  correspond to central (homogeneous)
    source distributions. The straight lines represent the fit to
    the data using the expresion in Eq. (\ref{eq:fwhm}).
    \label{fig:widthvsvelocity}} 
\end{figure*}

\begin{figure}
    \includegraphics[scale=0.35]{f3.eps}
\caption{Position of the peak maxima as a function of rotational
  velocity $V_{\rm max}$. Continuous (dashed) lines correspond to
  central (homogeneous) source distributions. A value of $x_{\rm
    max}=0$ indicates that line becomes single
  peaked. \label{fig:maximumsvsvelocity}}  
\end{figure}

There are three clear effects on the line's morphology as the
rotational veloctiy increases. First, the line
broadens; second, the double peaks reduce their intensity; and third,
the intensity at the line center rises. The two last effects give the
impresion that the double peaks are merger into one at high rotational
velocities, a result that is evident for the homogeneously distributed
sources as shown in the dashed lines of Figure \ref{fig:differentvelocities}.

We use the full width at half maximum (FWHM) to quantify the line
broadening. Figure \ref{fig:widthvsvelocity} shows how the FWHM increases
with rotational velocity both for the central an homogeneous source
distributions.  We parameterize the dependency of the linewidth as
\begin{equation}
 {\rm FWHM}^2 = {\rm FWHM}_{0}^2 + V_{\rm max}^2/\lambda^2,
\label{eq:fwhm}
\end{equation}
%
where FWHM$_{0}$ is the
velocity width in the static case and $\lambda$ is a positive scalar
to be determined as a fit to the data. With this test we want to know
to what extent the new velocity width can be expressed as a quadratic
sum of the two relevant velocities in the problem. We fit
simultaneously all the points corresponding to central and homogeneous
models to find $\lambda_{\rm c} = 3.9 \pm 3.17\times 10^{-17}$ and $\lambda_{\rm h}=
1.9\pm 3.07\times 10^{-10}$ respectively.  {\bf EL ERROR ES de $10^{-17}???$}

@@

\begin{figure}
    \includegraphics[scale=0.35]{f4.eps}
\caption{Logarithm of the average number of scatterings as function of
  rotational velocity. Continuous (dashed) lines represent an
  central (homogeneous) distribution of sources. \label{fig:Nscatt}}    
\end{figure}

\begin{figure*}
\begin{center}
  \includegraphics[scale=0.35]{f5_1.eps}
  \includegraphics[scale=0.35]{f5_2.eps}
  \includegraphics[scale=0.35]{f5_3.eps}
  \includegraphics[scale=0.35]{f5_4.eps}    
\end{center}
    \caption{2D histogram of $N_{\rm scatt}$ vs $x$. The upper (lower) panels
      show the homogeneous (central) source distribution. Left
      corresponds to the static case and right
      $V_{\rm max}=300\kms$. The color scale is logarithmic on the
      number of photons with given values of $N_{\rm scatt}$ and
      $x$. \label{fig:Nscatt2D}}   
\end{figure*}

The number of scatterings affects the escape frequency of a \ly photon.
In static cases, a large value of the optical depth correlates with a
high number of scatterings, increasing the probability of finding a
\ly photon far from the line's center. As a result the peak maxima
shifts from the center as the amount of neutral hydrogen
increases. This can be precisely quantifyied in the static slab
with central sources where the maxima position are related to
the optical depth and the temperature as $x_{m}=\pm 1.066(a\tau_{\rm
  H})^{1/3}$ \citep{Harrington73}. In the same model the average
number of scatterings depends only on the optical depth $\langle
N_{\rm  scatt}\rangle=1.612\tau_{\rm   H}$
\citep{Adams72,Harrington73}, in the case of homogeneously distributed
sources $\langle N_{\rm   scatt}\rangle=1.16\tau_{\rm   H}$
\citep{Harrington73}.

@@

\begin{table}
\begin{center}
\begin{tabular}{c ccc ccc}
\hline \hline
Sources & $\tau_{H}$ & & & $\ V_{\rm max}$& & \\
& &  &  & (\kms) & & \\ 
& & 0 & 50 & 100 &200 & 300\\ \hline 
Homogeneous & $10^{5}$& 0.263 & 0.266 &  0.309 &  0.357 &  0.370  \\
           & $10^{6}$ & 0.289 &  0.289 &   0.309 &  0.360 &  0.393 \\
         &$10^{7}$ &  0.227 &  0.229 &  0.231 &  0.254 &  0.281 \\
Central & $10^{5}$ &  0.096 & 0.097 & 0.096 &  0.096 & 0.096 \\
  &$10^{6}$ & 0.066 & 0.066 &  0.066 &  0.066 &  0.066 \\
 &$10^{7}$ & 0.015 &  0.016 & 0.016 & 0.016 & 0.015 \\
\hline
\end{tabular}
\caption{
 Escape fraction values for all dusty models. } 
\label{table:escape}
\end{center}
\end{table}

\begin{figure*}
\begin{center}
  \includegraphics[scale=0.35]{f6.eps}
  \includegraphics[scale=0.35]{f7.eps}
\end{center}
  \caption{Effects of rotation on the escape fraction. Left
    panel. Escape fraction as a function of rotational velocity. All
    these models have $\tau_{a}=1$. The continuous (dashed) lines
    correspond to central (homogeneous) models. Right panel. Escape
    fraction as a    function of the product $(a\tau_{\rm
      H})^{1/3}\tau_{a}$. The    analytic solution for the infinite
    slab is shown as a continuous    line. Different lines correspond
    to different rotational    velocities.     \label{fig:efvsv}} 
\end{figure*}

@@

Figure \ref{fig:efvsv} confirm these expectations. It shows the escape
fraction as a function of the rotational velocity. The curves for the
central source distribution stay flat, while for the homogeneous case
there is a clear rise with rotational velocity.  Rotation has a higher relative impact in
the models with low optical depth. For instance, in models with
$\tau_{\rm H}=10^5$, the static scape fraction is $0.26$ and increases
to $0.37$ for $V_{\rm max}=300$ \kms. Table \ref{table:escape} lists
all the values for the escape fraction.

In the right panel of Figure \ref{fig:efvsv} we put these results in the context of
the analytic solution for the infinite slab \citep{Neufeld90}. In
Neufeld's set-up the analytic solution depends uniquely on the product
$(a\tau_{\rm   H})^{1/3}\tau_{a}$, an approximation that is valid only
in the limit $a\tau_{\rm   H}\gg 1$. The dashed lines in the right panel of Figure
\ref{fig:efvsv} show the results for the different rotational
velocities. First of all we note that the escape fraction does not
increase from $\tau_{\rm H}=10^{5}$ to $\tau_{\rm H}=10^{6}$, a fact
that is explained since there is a transition from a opaque to an extremely opaque medium
which affect affects how \ly photons escape from the medium: single flight 
vs. single excursion. {\bf Que quiere decir eso de single flight y
  single excursion, porque explica eso la fraccion de escape?} 
 %we are in a regime where the condition for the analytic expectations
 %($a\tau_{h}\gg 1$) does not hold.

@@

\begin{figure}
\begin{center}
  \includegraphics[scale=0.35]{f8.eps}
\end{center}
\caption{Schematic representation of the geometry for the off-center
  emission. The large circle represents the gas distribution and the
    two smaller circles the off-centered \ly emitting regions. The angular
    velocity vector lies along $\hat{k}$ direction
    towards the reader. The vector $\hat{k}_{\rm out}$ represents the
    outgoing direction of an escaping \ly photon and the observer is
    placed along the $\hat{j}$ direction. The distribution of $\Theta$
    angles is used to study potential emission anisotropies. 
    \label{fig:OCspheres}}
\end{figure}

\begin{figure}
  \includegraphics[scale=0.35]{f9.eps}
  \caption{Histogram of $\mu\equiv\cos\Theta$ for the 
    central (dotted), homogeneous (dashed) and off-center (solid
    line) distributions. Deviations from a flat distribution signal the presence of
    spatial emission anisotropies. 
    \label{fig:muhisto}} 
\end{figure}

\begin{figure*}
\begin{center}
\includegraphics[scale=0.35]{f10.eps}
\end{center}
\caption{\ly profiles for the off-centered emission. Dashed/solid
  lines represents spheres centered at $-0,5R/+0.5R$ respectively.
   \label{fig:OC_profiles}} 
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[scale=0.35]{f11.eps}
\end{center}
\caption{Triple peaked profiles for three different off-center positions
  $C=0.25R$ (doted line), $C=0.5R$ (dashed line) \& $C=0.75R$ (solid
  line) and for different rotational velocities.  
   \label{fig:3p_profiles}} 
\end{figure*}

@@

An estimator for the anysotropy is represented by
%
\begin{equation}
F(\mu) = \frac{2\Delta N}{N\Delta \mu}, 
\end{equation} 
%
where $\mu=\cos\Theta$, $N$ is the total number of outgoing photons,
$\Delta N$ is the number of photons in a angular bin $\Delta
\Theta$. This definition satisfies the condition
$\int_{-1}^{1}F(\mu)d\mu/2=1$.  With this definitions, the isotropic
emission is characterized by a flat $F(\mu)$ distribution.

@@

We also explore a different source of anisotropy by studying the effect
of radiation sources located in off-centered spheres with respect to
the gas distribution. Figure \ref{fig:OCspheres} shows the two
spherical regions we define to test for this effect. Each has a radius
of $0.5R$, where $R$ is the radius of the gaseus sphere. With an
observer located along the $\hat{j}$ direction, we pick two possible
positions to center the spheres along the $\hat{k}$ axis: $\pm C\hat{k}$. In
this configuration the emitting sphere centered on $+C\hat{k}$ moves
away from the observer and the sphere centered on $-C\hat{j}$ moves
towards the observer.

In this case the $\mu$ distribution is not homogeneous as shown in
Figure \ref{fig:muhisto}. The off-center emision shows a variation of flux at
the levels of $10\% - 28\%$ in intensity with the angle $\Theta$. In general 
models with a larger $\tau_{H}$ present a higher variation in the flux.
This variation implies that that the line is more intense in
the plane perpendicular to the rotation axis.

We also explore the effects of off-centered emission in the line
morphology. Figure \ref{fig:OC_profiles} show sthe spectra for the two
off-centered emitting spheres located at $C=\pm0.5R$.  In this case we
only use photons that have outgoing directions aligned with the
observer i.e. $0.9 <\vert \mu \vert < 1.0$. We do not find any strong
asymmetry in the line morphology.

@@

The off-centered cases with hints of triple peaked emision motivate us
to explore this in more detail. The first condition is to increase the
number of photons in the radiative transfer calculation. In the line
profiles presented in the previous subsection, only $\sim 10^4$ photons
have the geometric conditions to build the line profile. We run a few
models with $N_{\gamma} = 10^6$ in order to improve the statistics and
explore smaller radii for the off-centered spheres.

@@

Furthermore, taking into account all the models that do not event present hints for
a triple peaked emission it would seem that the rotational velocity
and optical depth must balance each other in order to show this feature
feature. Namely, if the rotational velocity is close to half the HFWM
of the line, the triple peaks are more likely to occurr. This is the
same threshold that in the case of centered emission marks the
difference between double and single peaked lines.

@@

The presence of single peaked profiles has been associated to
inflow/outflow dynamics \citep{XXX}. In this paper we show that gas
bulk rotation can also be considered as a probable origin for that
behavior, provided that the observed single peak is highly
symmetric. Similarly, in the case of double peaked lines with a high
level of flux at the line center, rotation also deserves to be considered in
the pool of possible bulk flows responsible for that feature,
specially if the two peaks have similar intensities. The case of
triple peaked lines is another clear feature of gas rotation under the
additional condition of an off-centered emission.

@@

\item The line width increases with rotational velocity. The line with
  approximately following the functional for  ${\rm FWHM}^2 = {\rm FWHM}_{
    0}^2 + (V_{\rm max}/\lambda)^2$ where FWHM$_{0}$ indicates the line
  width for the static case. $\lambda$ is a constat determined from
  the radiative transfer results it is $\lambda_{\rm c}=3.9\pm$ and
  $\lambda_{\rm h}=1.9\pm$ for the central and homogeneous source
  distribution, respectively.

\item A single peaked line emerges for high rotational velocities in
  the case of homgeneously distributed sources. These cases occur when
  the rotational velocity is larger than hFWHM$_0$.

\item For central sources we find that the number of scatterings
does not significantly decrease with rotation. This is due to the 
fact that the vast majority of scatterings events are resonant, 
during which the mean free path is very short and the effect of gas
bulk rotation is not enough to affect \ly photons. In consequence, 
the escape fraction neither the position of the peaks depende on the
rotational velocity. However, there is a flux increase with rotational
velocity in the central parts of the line.

\item For homogeneous sources we find that the changes in the
line morphology are linked to the reduction in the average number of
scatterings. In the second part of \S \ref{sec:widthpeak} we show
that, as rotational velocity increases, a large fraction of photons
escape with less than $10$ scatterings. The avarage number of
scatterings decreases a $40\%$ compared to the static case. 
The reduction in the number of scattering decreases the probability
for a photon to move away from the line's center. This is also
translated into an increase of the escape fraction by a factor of
$\sim 2$ with respect to the static situation.

\item The spatial anisotropies in the emission are always at a level
  below the $10\%$ deviation from static models.

\item For off-centered emission triple peaked profiles can emerge
  under two conditions: the emission occurs close to th sphere's
  surface and the rotational velocity is half the FWHM in the static
  case.   
\end{itemize}

Comparing our results with recent observed LAEs we find that many 
morphological features such as high central line flux, single peak
profiles and multi-peaked profiles could be explained by gas bulk 
rotation present in these LAEs.

\section*{Acknowledgements}

>>>>

The codes follow the individual scatterings of \ly photons as it
travels through a 3D distribution of neutral Hydrogen. At each
scattering the frequency of the photon (in the laboratory frame) and
its direction of propagation change. This change in frequency is due
to the peculiar velocities of the Hydrogen atom that absorbed and
re-emitted the photon. If dust is present, the photon can interact
either with a Hydrogen atom or dust grain. In the case of a dust
interaction the photon can be either absorbed or scattered, this
probability is encoded in the dust albedo, $A$, which we chose to be
$1/2$. In order to obtain accurate values for the escape fraction of
photons in the presence of dust, we do not use any accelerating
mechanism in the radiative transfer. Once the photons escape the gas
distribution we store their direction at their direction of
propagation and frequency at their last scattering.

The photons are thus emitted in some region of the gas distribution
and follow a random walk in space and frequency until they escape the
gas distribution or are absorbed by a dust grain. The initialization
process for the \ly photons has to specify its position, frequency and
direction of propagation. In our case we select the initial frequency
to be exactly the \ly rest-frame frequency ($x=0$) and the direction of
propagation to be random following an flat probability distribution
over the sphere.

@@

\begin{figure*}
\begin{center}
  \includegraphics[width=0.8\textwidth]{f1.eps}
\end{center}
\caption{Shape of the \ly line for
    different velocities rotational velocities for spherical
    distributions in all the models. The continuous (dashed) line represents the 
    central (homogeneous) source distributions.
    \label{fig:differentvelocities}}  
\end{figure*}

@@

In the next subsections we characterize the line morphology by the
full width at half maximum (FWHM) and the location of the peak
maxima. In order to interpret the morphological changes in the line we
also report the median number of scatter for each \ly photon in the
simulation. For the models where dust is included we measure the 
escape fraction as a function of rotational velocity. Finally, we make
an estimate of the anisotropic emission of the models in comparison
with static spheres.

@@

\begin{figure*}
\begin{center}
  \includegraphics[width=0.8\textwidth]{f2.eps}
\end{center}
  \caption{FWHM for the non-dusty models as a function of
    rotational velocity $V_{\rm max}$. The left panel shows the
    results in velocity units while the right panel normalizes
    the data by the FWHM in the static case. 
    Continuous (dashed) lines  correspond to central (homogeneous)
    source distributions. The straight lines represent the fit to
    the data using the expression in Eq. (\ref{eq:fwhm}).
    \label{fig:widthvsvelocity}} 
\end{figure*}

\begin{figure}
\begin{center}
  \includegraphics[width=0.40\textwidth]{f3.eps}
\end{center}
\caption{Position of the peak maxima as a function of rotational
  velocity $V_{\rm max}$. Continuous (dashed) lines correspond to
  central (homogeneous) source distributions. A value of $x_{\rm
    max}=0$ indicates that line becomes single
  peaked. \label{fig:maximumsvsvelocity}}  
\end{figure}

There are three clear effects on the line's morphology as the
rotational velocity increases. First, the line
broadens; second, the double peaks reduce their intensity; and third,
the intensity at the line center rises. The two last effects give the
impression that the double peaks are merger into one at high rotational
velocities, a result that is evident for the homogeneously distributed
sources as shown in the dashed lines of Figure \ref{fig:differentvelocities}.

We use the full width at half maximum (FWHM) to quantify the line
broadening. Figure \ref{fig:widthvsvelocity} shows how the FWHM increases
with rotational velocity both for the central an homogeneous source
distributions.  We parameterize the dependency of the line width as
\begin{equation}
 {\rm FWHM}^2 = {\rm FWHM}_{0}^2 + V_{\rm max}^2/\lambda^2,
\label{eq:fwhm}
\end{equation}
%
where FWHM$_{0}$ is the
velocity width in the static case and $\lambda$ is a positive scalar
to be determined as a fit to the data. With this test we want to know
to what extent the new velocity width can be expressed as a quadratic
sum of the two relevant velocities in the problem. We fit
simultaneously all the points corresponding to central and homogeneous
models to find $\lambda_{\rm c} = 3.9 \pm 3.17\times 10^{-17}$ and $\lambda_{\rm h}=
1.9\pm 3.07\times 10^{-10}$ respectively.  {\bf EL ERROR ES de $10^{-17}???$}

@@

\begin{figure}
\begin{center}
    \includegraphics[width=0.40\textwidth]{f4.eps}
\end{center}
\caption{Logarithm of the average number of scatterings as function of
  rotational velocity. Continuous (dashed) lines represent an
  central (homogeneous) distribution of sources. \label{fig:Nscatt}}    
\end{figure}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.45\textwidth]{f5_1.eps}
  \includegraphics[width=0.45\textwidth]{f5_2.eps}
  \includegraphics[width=0.45\textwidth]{f5_3.eps}
  \includegraphics[width=0.45\textwidth]{f5_4.eps}    
\end{center}
    \caption{2D histogram of $N_{\rm scatt}$ vs $x$. The upper (lower) panels
      show the homogeneous (central) source distribution. Left
      corresponds to the static case and right
      $V_{\rm max}=300\kms$. The color scale is logarithmic on the
      number of photons with given values of $N_{\rm scatt}$ and
      $x$. \label{fig:Nscatt2D}}   
\end{figure*}

The number of scatterings affects the escape frequency of a \ly photon.
In static cases, a large value of the optical depth correlates with a
high number of scatterings, increasing the probability of finding a
\ly photon far from the line's center. As a result the peak maxima
shifts from the center as the amount of neutral hydrogen
increases. This can be precisely quantified in the static slab
with central sources where the maxima position are related to
the optical depth and the temperature as $x_{m}=\pm 1.066(a\tau_{\rm
  H})^{1/3}$ \citep{Harrington73}. In the same model the average
number of scatterings depends only on the optical depth $\langle
N_{\rm  scatt}\rangle=1.612\tau_{\rm   H}$
\citep{Adams72,Harrington73}, in the case of homogeneously distributed
sources $\langle N_{\rm   scatt}\rangle=1.16\tau_{\rm   H}$
\citep{Harrington73}.

@@

\begin{table}
\begin{center}
\begin{tabular}{c ccc ccc}
\hline \hline
Source & $\tau_{H}$ & & & $\ V_{\rm max}$& & \\
Distribution& &  &  & (\kms) & & \\ 
& & 0 & 50 & 100 &200 & 300\\ \hline 
Homogeneous & $10^{5}$& 0.263 & 0.266 &  0.309 &  0.357 &  0.370  \\
           & $10^{6}$ & 0.289 &  0.289 &   0.309 &  0.360 &  0.393 \\
         &$10^{7}$ &  0.227 &  0.229 &  0.231 &  0.254 &  0.281 \\
Central & $10^{5}$ &  0.096 & 0.097 & 0.096 &  0.096 & 0.096 \\
  &$10^{6}$ & 0.066 & 0.066 &  0.066 &  0.066 &  0.066 \\
 &$10^{7}$ & 0.015 &  0.016 & 0.016 & 0.016 & 0.015 \\
\hline
\end{tabular}
\caption{
 Escape fraction values for all dusty models. } 
\label{table:escape}
\end{center}
\end{table}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.40\textwidth]{f6.eps}
  \includegraphics[width=0.40\textwidth]{f7.eps}
\end{center}
  \caption{Effects of rotation on the escape fraction. Left
    panel. Escape fraction as a function of rotational velocity. All
    these models have $\tau_{a}=1$. The continuous (dashed) lines
    correspond to central (homogeneous) models. Right panel. Escape
    fraction as a    function of the product $(a\tau_{\rm
      H})^{1/3}\tau_{a}$. The    analytic solution for the infinite
    slab is shown as a continuous    line. Different lines correspond
    to different rotational    velocities.     \label{fig:efvsv}} 
\end{figure*}

@@

Figure \ref{fig:efvsv} confirm these expectations. It shows the escape
fraction as a function of the rotational velocity. The curves for the
central source distribution stay flat, while for the homogeneous case
there is a clear rise with rotational velocity.  Rotation has a higher relative impact in
the models with low optical depth. For instance, in models with
$\tau_{\rm H}=10^5$, the static escape fraction is $0.26$ and increases
to $0.37$ for $V_{\rm max}=300$ \kms. Table \ref{table:escape} lists
all the values for the escape fraction.

In the right panel of Figure \ref{fig:efvsv} we put these results in the context of
the analytic solution for the infinite slab \citep{Neufeld90}. In
Neufeld's set-up the analytic solution depends uniquely on the product
$(a\tau_{\rm   H})^{1/3}\tau_{a}$, an approximation that is valid only
in the limit $a\tau_{\rm   H}\gg 1$. The dashed lines in the right panel of Figure
\ref{fig:efvsv} show the results for the different rotational
velocities. First of all we note that the escape fraction does not
increase from $\tau_{\rm H}=10^{5}$ to $\tau_{\rm H}=10^{6}$, a fact
that is explained since there is a transition from a opaque to an extremely opaque medium
which affect affects how \ly photons escape from the medium: single flight 
vs. single excursion. {\bf Que quiere decir eso de single flight y
  single excursion, porque explica eso la fraccion de escape?}

@@

\begin{figure}
\begin{center}
  \includegraphics[width=0.40\textwidth]{f8.eps}
\end{center}
\caption{Schematic representation of the geometry for the off-center
  emission. The large circle represents the gas distribution and the
    two smaller circles the off-centered \ly emitting regions. The angular
    velocity vector lies along $\hat{k}$ direction
    towards the reader. The vector $\hat{k}_{\rm out}$ represents the
    outgoing direction of an escaping \ly photon and the observer is
    placed along the $\hat{j}$ direction. The distribution of $\Theta$
    angles is used to study potential emission anisotropy. 
    \label{fig:OCspheres}}
\end{figure}

\begin{figure}
  \includegraphics[width=0.42\textwidth]{f9.eps}
  \caption{Histogram of $\mu\equiv\cos\Theta$ for the 
    central (dotted), homogeneous (dashed) and off-center (solid
    line) distributions. Deviations from a flat distribution signal the presence of
    spatial emission anisotropy. 
    \label{fig:muhisto}} 
\end{figure}

\begin{figure*}
\begin{center}
\includegraphics[width=0.90\textwidth]{f10.eps}
\end{center}
\caption{\ly profiles for the off-centered emission. Dashed/solid
  lines represents spheres centered at $-0,5R/+0.5R$ respectively.
   \label{fig:OC_profiles}} 
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=0.90\textwidth]{f11.eps}
\end{center}
\caption{Triple peaked profiles for three different off-center positions
  $C=0.25R$ (doted line), $C=0.5R$ (dashed line) \& $C=0.75R$ (solid
  line) and for different rotational velocities.  
   \label{fig:3p_profiles}} 
\end{figure*}

@@

An estimator for the anisotropy is represented by
%
\begin{equation}
F(\mu) = \frac{2\Delta N}{N\Delta \mu}, 
\end{equation} 
%
where $\mu=\cos\Theta$, $N$ is the total number of outgoing photons,
$\Delta N$ is the number of photons in a angular bin $\Delta
\Theta$. This definition satisfies the condition
$\int_{-1}^{1}F(\mu)d\mu/2=1$.  With this definitions, the isotropic
emission is characterized by a flat $F(\mu)$ distribution.

@@

We also explore a different source of anisotropy by studying the effect
of radiation sources located in off-centered spheres with respect to
the gas distribution. Figure \ref{fig:OCspheres} shows the two
spherical regions we define to test for this effect. Each has a radius
of $0.5R$, where $R$ is the radius of the gaseous sphere. With an
observer located along the $\hat{j}$ direction, we pick two possible
positions to center the spheres along the $\hat{k}$ axis: $\pm C\hat{k}$. In
this configuration the emitting sphere centered on $+C\hat{k}$ moves
away from the observer and the sphere centered on $-C\hat{j}$ moves
towards the observer.

In this case the $\mu$ distribution is not homogeneous as shown in
Figure \ref{fig:muhisto}. The off-center emission shows a variation of flux at
the levels of $10\% - 28\%$ in intensity with the angle $\Theta$. In general 
models with a larger $\tau_{H}$ present a higher variation in the flux.
This variation implies that that the line is more intense in
the plane perpendicular to the rotation axis.

We also explore the effects of off-centered emission in the line
morphology. Figure \ref{fig:OC_profiles} show the spectra for the two
off-centered emitting spheres located at $C=\pm0.5R$.  In this case we
only use photons that have outgoing directions aligned with the
observer i.e. $0.9 <\vert \mu \vert < 1.0$. We do not find any strong
asymmetry in the line morphology.

@@

The off-centered cases with hints of triple peaked emission motivate us
to explore this in more detail. The first condition is to increase the
number of photons in the radiative transfer calculation. In the line
profiles presented in the previous subsection, only $\sim 10^4$ photons
have the geometric conditions to build the line profile. We run a few
models with $N_{\gamma} = 10^6$ in order to improve the statistics and
explore smaller radii for the off-centered spheres.

@@

Furthermore, taking into account all the models that do not event present hints for
a triple peaked emission it would seem that the rotational velocity
and optical depth must balance each other in order to show this feature
feature. Namely, if the rotational velocity is close to half the HFWM
of the line, the triple peaks are more likely to occur. This is the
same threshold that in the case of centered emission marks the
difference between double and single peaked lines.

@@

The presence of single peaked profiles has been associated to
inflow/outflow dynamics \citep{Verhamme06,DijkstraKramer}. In this
paper we show that gas bulk rotation can also be considered as a
probable origin for that behavior, provided that the observed single peak is highly
symmetric. Similarly, in the case of double peaked lines with a high
level of flux at the line center, rotation also deserves to be considered in
the pool of possible bulk flows responsible for that feature,
specially if the two peaks have similar intensities. The case of
triple peaked lines is another clear feature of gas rotation under the
additional condition of an off-centered emission.

@@

\item The line width increases with rotational velocity. The line with
  approximately following the functional for  ${\rm FWHM}^2 = {\rm FWHM}_{
    0}^2 + (V_{\rm max}/\lambda)^2$ where FWHM$_{0}$ indicates the line
  width for the static case. $\lambda$ is a constant determined from
  the radiative transfer results it is $\lambda_{\rm c}=3.9\pm$ and
  $\lambda_{\rm h}=1.9\pm$ for the central and homogeneous source
  distribution, respectively.

\item A single peaked line emerges for high rotational velocities in
  the case of homogeneously distributed sources. These cases occur when
  the rotational velocity is larger than FWHM$_0$.

\item For central sources we find that the number of scatterings
does not significantly decrease with rotation. This is due to the 
fact that the vast majority of scatterings events are resonant, 
during which the mean free path is very short and the effect of gas
bulk rotation is not enough to affect \ly photons. In consequence, 
the escape fraction neither the position of the peaks depends on the
rotational velocity. However, there is a flux increase with rotational
velocity in the central parts of the line.

\item For homogeneous sources we find that the changes in the
line morphology are linked to the reduction in the average number of
scatterings. In the second part of \S \ref{sec:widthpeak} we show
that, as rotational velocity increases, a large fraction of photons
escape with less than $10$ scatterings. The average number of
scatterings decreases a $40\%$ compared to the static case. 
The reduction in the number of scattering decreases the probability
for a photon to move away from the line's center. This is also
translated into an increase of the escape fraction by a factor of
$\sim 2$ with respect to the static situation.

\item The spatial anisotropy in the emission is always below the
  $10\%$ level deviation from static models.

\item For off-centered emission triple peaked profiles can emerge
  under two conditions: the emission occurs close to the sphere's
  surface and the rotational velocity is half the FWHM in the static
  case.   
\end{itemize}

Comparing our results with recent observed LAEs we find that many 
morphological features such as high central line flux, single peak
profiles and multiple-peaked profiles could be explained by gas bulk 
rotation present in these LAEs.

\section*{Acknowledgments}
